Fish do not need Visas: the transboundary nature of the world’s exploited marine fish species (79/96 characters)

### 2020-01-02 Manuscript_Draft.Rmd
Reduced text lenght to accomodate the 2500 words (With Refs.). Currently under 2000 (Without Refs.)

- **Potential Reviewers (J)**  
 -- Gakushi Ishimura,  
 -- Kathleen Miller, National Center for Atmospheric Research, Boulder, Colorado, USA  
 -- Malin Pinsky, Rutgers University, New Brunswick, New Jersey, USA  
 -- Steven D. Gaines, University of California, Santa Barbara, Santa Barbara, California, USA

@@

Distributions of marine species around the world are not bounded by human-made boundaries; rather shaped by contemporary biotic and abiotic factors, as well as their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. Thus, a species can be cosmopolitan and have a broader distribution like the spiny dogfish (*Squalus acanthias*), or endemic like the masu salmon (*Oncorhynchus masou*) who's distribution is limited to the Japanese coast [@Nelson:2016kk]. However, management of living marine resources is predicated on the definition of “stocks” that are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, established political boundaries across the distribution of many species, creating shared species between nations. Theories and empirical experience demonstrate fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@Bailey:2010ga]. For example, environmental fluctuation leading to changes in species distribution of a shared fishery have led to conflict because of changes in the proportion of captures between countries [@Munro:2002uf], or pressure in the quota allocation due to new participants wanting to join the fishery [@Spijkers:2017ij]. Moreover, climate change, which is shifting species distributions, is likely to exacerbate these types of conflict [@Pinsky:2018cb] and presents further challenges for fisheries management [@Miller:2013iv].

Forty years after the formal adoption of UNCLOS, accurate estimates of the number of exploited marine species that are shared by 2 or more neighboring coastal nations (hereafter called transboundary species [@Munro:2002uf]) in the world is still unclear. Caddy in 1997 estimated that there could be up to 1,500 transboundary fish species in the world. However, such estimation lacked a proper assessment due to limited information on the distribution of species at the time [@Caddy:1997ue]. More recently, Teh and Sumaila (2010) estimated that 206 species can be considered shared, accounting for a total catch of 34.2 million tones and a global landed value of USD 30.7 trillion. Such analysis was based on a literature review of shared species and did not include a mechanistic way of determining the transboundary nature of the species [@Teh:2015gd]. In here, we overlay the known distribution of **968** commercially valuable marine fish species and the EEZ of 153 coastal countries to identify those that are shared between neighboring EEZs and determine their contribution to the world's fisheries catch and revenue.

We developed a four-criteria methodology to determine if a species was transboundary (see Methods). The first criteria consisted in only considering shared species between neighboring EEZs, rather than the specie's extended distribution. For the second criteria we considered a species to be present in a $0.5^o$ x $5^o$ grid cell if all data sources predicted positive occurrence. Current species distributions were predicted from occurrence records obtained from publicly available observational databases of the world’s marine species and two structurally distinct species distribution models [@Asch:2018ca; Reygondeau:2019uh; @Zeller2016]. The third criteria determined whether the species were exploited at the grid cell using spatially explicit catch data [@Zeller2016]. A species was kept in a grid cell if both predicted species occurrences and catch data were positive. Finally, the last criteria determined a spatial threshold to confirm that a species occurs in both neighboring EEZs. Once transboundary species were identified, we estimated their contribution to fishing landings and revenues using the *Sea Around Us* database [@Zeller2016] and categorized them according to catch trends [@Kleisner:2013gh]. We carry out the analysis at the species per EEZ level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species [@Teh:2015gd]. Despite loosing detail-specific (e.g. a species could have multiple stocks within an EEZ),  many fisheries within a country or EEZ are managed at the species level, and recent research has suggested strong connectivity between stocks thousand of kilometers away [@Ramesh:2019va].

@@

We identified **633** transboundary species in the world, more than double the number previously estimated [@Teh:2015gd], accounting for **65.4%** of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared species within EEZs landed an average of **69.2** million tonnes, representing **64%** of the EEZ-global catch, and generating USD **79,745** million in fishing revenue. This is equivalent to **49.8%** of the global revenue of the same period. These values are higher than previously estimated reinforcing the importance of these species at the global level [@Teh:2015gd].

@@

![**Fig. 1 Number of transboundary species and they contribution to fishing catch and revenue.** A) The number of species is represented in the EEZ while revenue from these species fished within the EEZ is represented in the land polygon. B) Regional revenue (left) and catch (right) contribution of tranbsoundary species. Points = mean +- sd.](./Figures/Fig1.png)

Transboundary species are particularly important for Northern America (mean = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (mean = 2,827 $\pm$ = 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). The large variation in revenue from these regions is driven by China (USD 6,279 million), in E. Asia and the US (USD 9,605 million) in N. America who contribute 55% and 83% to the region's revenue, respectively. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with more revenue from transboundary species. These five nations are responsible for 30% of the global revenue from transboundary species. Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead the production of transboundary species harvesting over 5,000 thousand tones (**Supplemental figure Trans_Catch_MAP**). Peru's catches come mainly from Peruvian anchoveta (*Engraulis ringens*), shared with Chile, accounting for 79% of the transboundary species production. Similarly, Russia shares the Alaskan polllock (*Theragra chalcogramma*) fishery with both Japan and the US. Despite this, catches are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@FAO:2018KK] getting around 1/3 of its capture from transboundary species, followde by Japan (4.2 million tonnes) and South Korea (2.3 million tonnes).

Effective management of shared species is a challenging process as it often requires countries to agree on mutual targets in order to achieve sustainable harvest [@Munro:2002uf]. We found significant differences in the exploitation rate of transboundary species (one-way ANOVA *F*(5,446), *f* = 324.5, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05 for all) with most species fished at levels below 50%, sometimes lower than 10%, of their historical maximum peak (Fig.2). This is the case of coastal nations in all continents, but Oceania, with no significant difference between continents (one-way ANOVA *F*(5,443), *f* = 1.24, *p* = 0.289, $\alpha$ = 0.05), suggesting that the current global exploitation rate of transboundary species could be over the maximum sustainable yield [@Kleisner:2011wn; @Pauly:2013gs]. While these catch trends do not directly reflect stock status - e.g. other factors like markets, fishing policies or extreme events also influence catch volume [@Pauly:2013gs; @Branch:2008kk] - they could resemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. Countries that are today harvesting transboundary species at this rate (Category C) should revise their management policies to make sure these align with sustainable harvest of shared resources [@Munro:2004ug].

@@

![**Number of transboundary species by catch category, country, and continent. Category A, fishing between 10%-50% of max catch peak after a drop larger than 10% of max peak; Category B, fishing over 50% of max catch peak; Category C, fishing is between 10-50% of max catch peak or less than 10%**](./Figures/Fig2.png)

Empirical analysis suggest that in most cases, management of transboundary species will have better outcomes in terms of fish yield when nations cooperate [@Bailey:2010ga]. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*) [@Diekert:2010gp; @Eide:2013dha], and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Lack of collaboration in shared stocks may threaten stock sustainability [@Clark:1980js; @Nguyen:2018dk], reduce the profitability potential of the fishery [@Merino:2007jz], although cooperation and non-cooperation can be more complex [@Jensen:2015cf]. Collaboration can also reduce conflict between nations over shared resources. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts like Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations to compensate losses [@Munro:2002uf]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for Atlantic mackerel (*Scomber scombrus*) [@Spijkers:2017ij], a species that is transboundary and contributes **698.3** thousand tones to landings and USD **968** million in revenue to the countries involved in the dispute. In addition to past and present conflicts, climate change could exacerbate conflicts over transboundary species as climate change shifts the distribution of marine species around the globe [@Pinsky:2018cb].

@@

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], many of which are highly migratory pelagic-oceanic species like skipjack tuna (*Katsuwonus pelamis*), Common dolphinfish (*Coryphaena hippurus*), and tunas (*Thunnus sp.*). This is the only group whoes EEZ range (e.g. the number of EEZs where the species occur as transboundary) significantly differs from the others (one-way ANOVA *F*(5,109), *f* = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species (Fig. 3).

@@

![**Number of countries shared by transboundary species. Showing only species that share > 20 countries**](./Figures/Fig3.png)

Due to their large migration patterns and presence in areas beyond national jurisdiction, highly migraroty pelagic fish species are cataloged as straddling and many of them globally managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species, most of them fished in areas beyond national jurisdiction [@CullisSuzuki:2010fi]. We estimate that 81.8% of these species are actually shared within neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighboring RFMOs (**Supplemental figure RFMO_MAP**). The effectiveness of RFMOs responsible for managing high migratory stocks has been previously questioned [@CullisSuzuki:2010fi; Lodge:2007bt]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 13 species (40 stocks) suggests that most tuna and billfish stocks are sustainably harvested ($B/B_{MSY} > 1$ and $F/F_{MSY} < 1$) with only three stocks currently overfished [@Pons:2016ep]. Within national waters, we estimated that the catch of transboundary pelagic-oceanic species comes mainly from five Island Nations in the South Pacific from which three (Kiribalti, Solomon and Papua New Guinea) are members of the Pacific Nauru Agreement (Fig 3B). The parties to the Nauru Agreement sustainably manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

@@

# Materials and Methods

We determine whether or not a species can be considered transboundary. The method is based on a series of criteria to be met and the overlay of species distributions and the spatial boundaries of the world’s EEZs. For the current analysis we work at the species level and adopt the Food and consider transboundary species as those that occur within the Exclusive Economic Zone (EEZ) of neighboring countries [@Munro:2002uf]. All data was scaled to 0.5 degree latitude x 0.5 degree longitude grid cells (*n* = ~180,000 grid cells).

## Databases on species geographic distributions

To determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs we first extracted all the commercial marine fish or invertebrate species from the *Sea Around Us* database (http://www.seaaroundus.org and determined their current distributions. For this, we used four data sources of species-distributions: (*i*) observational data, (*ii*) an Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (**Table x**). Each source represent a different method of estimating the distribution of a given species and thus, provides a more robust result. Only commercial fished species with data from all four sources were included in the analysis. The final dataset comprised **XXX** species, **XX**% of the reported marine species in the SAU database and **XX**% of the reported taxa of the FAO (**See S1_Data**).

### *Occurence data*

The occurrence data was collected from five publicly available repositories; Fishbase (http://fishbase.org), the Global Biodiversity Information Facility (GBIF; https://www.gbif.org/), the Ocean Biogeographic Information System (OBIS; https://obis.org/),  the Intergovernmental Oceanographic Comission (IOC; http://ioc-unesco.org), and the International Union for Conservation of Nature (IUCN; https://www.iucn.org/technical-documents/spatial-data) [@Reygondeau:2019uh].

### *Distribution models*

In addition to the occurrence data we use two different methods to estimate species distributions, hereafter referred as ENM-Nereus and SDM-SAU. Although they use the same data, the models are structurally different complementing each other and providing robustness to the results.

The ENM-Nereus consists in a multimodel approach based on a Bioclim and a Bososted Regression Tree model [@Thuiller:2009gp], a Maxent model [@Phillips:2006ff], and a Non-Parametric Probabilistic Ecological Niche Model [@Beaugrand:2011fd]. Environmental variables utilized in the models included sea surface temperature, surface pH, surface oxygen concentration, and vertically integrated (0–100 m) net primary production (NPP) [@Asch:2018ca]. The ENM-Nereus employed all of the observational data-sources previously mentioned. Global environmental conditions were obtained from 1970 to 2000 and average from three Earth System Models developed by the Geophysical Fluid Dynamics Laboratory (GFDL- https://www.gfdl.noaa.gov/earth-system-model/), the Institute Pierre Simon Laplace (IPSL- www.icmc.ipsl.fr/), and the Max Planck Institute for Meteorology (MPI- www.mpimet.mpg.de/en/science/models/). See [@Asch:2018ca; @Reygondeau:2019uh] for model details.

The SDM-SAU model follows a five-steps process based on species-specific life history information, rather than environmental variables [@Close:2006ux; @Palomares:wg]. For each commercial fish species, the model first uses the FAO major fishing areas and countries EEZs to determine a broad distribution. It then uses life history information to delimit its range within the FAO fishing area (e.g. thermal preference, depth limits). The range is delimited even further by expert-review polygons and compared with that of AquaMaps [@Kaschner:2016tl], OBIS and GBIF occurrence data. The model then determines a species habitat preference based on the assumptions that the relative abundance of a species is determined by the number of habitats in a grid cell and the distance of the species to each habitat, as well as the importance of the habitat to the species size. Finally, the species equatorial submergence (e.g. the latitudinal region where a species is not seen in between poles) is estimated for each species. See [@Close:2006ux; @Palomares:wg] for model details.

### *Catch data*

The previous models combine observational data with a series of biotic and abiotic information to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g. ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - i.e., catches  based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in the FAO data available for each country. Catches are also spatially allocated [@Zeller2016]. Thus, we used the Sea Around Us catch reconstruction database from 2005 to 2014 (last ten years of data) as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs.

@@

## Determine if a species is transboundary

We developed a four-criteria method for determining whether or not a species was transboundary. Only species that meet all criteria were considered as transboundary species and the analysis was done only within the EEZs of coastal states. In some cases, criteria also work as indexes to measure the uncertainty in the analysis as they provided continuous index (e.g. from 0 to 1) rather than a discrete, yes or no, index . All of the analysis was done in the statistical software *R version 3.5.2 (2018-12-20)* with the packages *data.table*[@Packagedatatable:2019uh], *janitor*[@PackagejanitorSim:2018ur], *wesanderson*[@Packagewesanderson:2018vn], *rfishbase*[@PackagerfishbaseR:2019th], *R.matlab*[@PackageRmatlabRe:2018un], *sf*[@PackagesfSimpleF:2018vp], *sp*[@PackagespClasses:2019tn], *tidiverse*[@PackagetidyverseE:2017vq], *tidytext*[@PackagetidytextTe:2019vk], and *zoo*[@PackagezooSInfr:2019tm]. All code is available at https://github.com/jepa/FishForVisa.

### Criteria 1; Neighboring EEZs

As previously mentioned, we define transboundary species as those marine species that happen within the EEZs of two or more neighboring countries. Hence, the first criteria was that the species would only be transboundary between two neighboring countries, regardless of the species extended distribution. This way, a species such as Atlantic cod (*Gadus morhua*) distributed along the north Atlantic, was only considered transboundary between each of the neighboring nations covering its distribution, rather than the region as a whole. We assume that this would reduce the differences between species and stock as is more likely that neighboring species belong to the same stock, rather than, lets say, Cod in Norway and Canada. However, the analysis here presented was kept at the species level and stocks within countries were not considered.

We define the EEZ boundaries using the SAU shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) that sub-divides EEZs by regions (e.g. Mexico Pacific and Mexico Atlantic) and determine the intersections between polygons using the *sf*-*R* package. When estimating transboundary species, we filter out those shared by the EEZs sub-regions, and when aggregating by country we only counted the species once, if it appeared in more than one sub-region. Species that were present in EEZs that were non-continental territories neighboring other countries were kept (e.g. Argentina and Falkland Islands), but removed when the non-continental territory belonged to the same nation (e.g. Brazil and Fernando de Noronha).

### Criteria 2; Data agreement

>>>>

- **Potential Reviewers (J)**  
 -- Gakushi Ishimura,  
 -- Kathleen Miller, National Center for Atmospheric Research, Boulder, Colorado, USA  
 -- Malin Pinsky, Rutgers University, New Brunswick, New Jersey, USA  
 -- Megan Baily, 
 -- Steven D. Gaines, University of California, Santa Barbara, Santa Barbara, California, USA

@@

Distributions of marine species around the world are not restricted by human-made boundaries; rather shaped by contemporary biotic and abiotic factors, as well as their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. Thus, a species can have a broader distribution (cosmopolitan), or be endemic [@Nelson:2016kk]. However, fisheries management is predicated on the definition of “stocks” that are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, established political boundaries across the distribution of many species, creating shared species between nations. Theories and empirical experience demonstrate fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@Munro:2002uf; @Bailey:2010ga]. For example, climate-driven changes in species distributions have led to conflict between nations due to changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate these types of conflict [@Pinsky:2018cb] and presents further challenges for fisheries management [@Miller:2013iv].

Forty years after the formal adoption of UNCLOS, accurate estimates of the number of exploited marine species that are shared by neighboring nations (hereafter called transboundary species [@Munro:2002uf]) is still unclear due to data limitations [@Caddy:1997ue], and methodological approaches [@Teh:2015gd]. In here, we aimed to estimate the number of exploited marine fish species shared by neighbouring countries and determine their contribution to global catch and fisheries revenue.

We overlay the known distribution of **968** commercially valuable marine species and the EEZ of 153 coastal countries and developed a four-criteria methodology to determine if a species was transboundary. The first criteria consisted in only considering shared species between neighboring EEZs, rather than the specie's extended distribution. For the second criteria we considered a species to be present in a grid cell if all data sources predicted positive occurrence. The third criteria determined whether the species were exploited at the grid cell using spatially explicit catch data. A species was kept in a grid cell if both predicted species occurrences and catch data were positive. Finally, the last criteria determined a spatial threshold to confirm that a species occurs in both neighboring EEZs. Once transboundary species were identified, we estimated their contribution to fishing landings and revenues and categorized them according to catch trends. We carry out the analysis at the species per EEZ level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species [@Teh:2015gd]. Despite loosing detail-specific (e.g. a species could have multiple stocks within an EEZ), many fisheries within a country or EEZ are managed at the species level, and recent research has suggested strong connectivity between stocks thousand of kilometers away [@Ramesh:2019va].

@@

We identified **633** transboundary exploited species in the world, more than double the number previously estimated [@Teh:2015gd], accounting for **65.4%** of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared species within EEZs landed an average of **69.2** million tonnes, representing **64%** of the EEZ-global catch, and generating USD **79,745** million in fishing revenue. This is equivalent to **49.8%** of the global revenue of the same period. These values are higher than previously estimated [@Teh:2015gd], reinforcing the importance of these species at the global level.

@@

![Fig. 1. Number of transboundary species and they contribution to fishing catch and revenue. A) The number of species and revenue are represented in the EEZ and land polygon, respectively. B) Contribution of tranbsoundary species to regional revenue (left) and catch (right). Points = mean +- sd.](./Figures/Fig1.png)

Transboundary species are particularly important for Northern America (mean = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (mean = 2,827 $\pm$ = 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). The large variation in revenue from these regions are driven by China (USD 6,279 million) in E. Asia and the US (USD 9,605 million) in N. America who contribute 55% and 83% to the region's revenue, respectively. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with more revenue from transboundary species. These five nations are responsible for 30% of the global revenue from transboundary species. Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead the production of transboundary species harvesting over 5,000 thousand tones (**Supplemental figure Trans_Catch_MAP**). Peru's catches come mainly from Peruvian anchoveta (*Engraulis ringens*), shared with Chile, accounting for 79% of the transboundary species production. Similarly, Russia shares the Alaskan polllock (*Theragra chalcogramma*) fishery with both Japan and the US. Despite this, catches are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@FAO:2018KK] getting around 1/3 of its capture from transboundary species, followed by Japan (4.2 million tonnes) and South Korea (2.3 million tonnes).

Harvest rates and fisheries management are key to achieve the sustainable development goals [@SDGs]. We found significant differences in the exploitation rate of transboundary species (one-way ANOVA *F*(5,446), *f* = 324.5, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05 for all) with most species fished at levels below 50%, sometimes lower than 10%, of their historical maximum peak (Fig.2). This is the case of coastal nations in all continents, but Oceania, with no significant difference between continents (one-way ANOVA *F*(5,443), *f* = 1.24, *p* = 0.289, $\alpha$ = 0.05). While these catch trends do not directly reflect stock status - e.g. other factors like markets, fishing policies or extreme events also influence catch volume [@Pauly:2013gs; @Branch:2008kk] - they could resemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. Thus suggesting that the current global exploitation rate of transboundary species could be over the maximum sustainable yield [@Kleisner:2011wn; @Pauly:2013gs]. Corroborating the status of species harvested at this rate (Category C) and considering best practices in managing shared resources [@Munro:2004ug; @Lodge:2007bt] could provide crucial to the sustainability of transboundary species.

@@

![Fig. 2. Number of transboundary species by catch category. Category A, fishing between 10%-50% of max catch peak after a drop larger than 10% of max peak; Category B, fishing over 50% of max catch peak; Category C, fishing less than 50% of max peak](./Figures/Fig2.png)

Empirical analysis suggest that in most cases, management of transboundary species will have better outcomes in terms of fish yield when nations cooperate [@Bailey:2010ga]. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*) [@Diekert:2010gp; @Eide:2013dha], and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Lack of collaboration in shared stocks may threaten stock sustainability [@Clark:1980js; @Nguyen:2018dk], reduce the profitability potential of the fishery [@Merino:2007jz], although cooperation and non-cooperation can be a more complex process [@Jensen:2015cf]. Collaboration can also reduce conflict between nations over shared resources. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts like Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations to compensate losses [@Munro:2002uf]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for Atlantic mackerel (*Scomber scombrus*) [@Spijkers:2017ij], a species that is transboundary and contributes **698.3** thousand tones to landings and USD **968** million in revenue to the countries involved in the dispute. In addition to past and present conflicts, climate change could exacerbate conflicts over transboundary species as climate change shifts the distribution of marine species around the globe requiring flexible agreements that accommodate future uncertainties [@Pinsky:2018cb].

@@

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], many of which are highly migratory pelagic-oceanic species like skipjack tuna (*Katsuwonus pelamis*), Common dolphinfish (*Coryphaena hippurus*), and tunas (*Thunnus sp.*). Pelagic-oceanic species are the only group whoes EEZ range (e.g. the number of EEZs where the species occur as transboundary) significantly differs from the others (one-way ANOVA *F*(5,109), *f* = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species (Fig. 3).

@@

![Fig. 3. EEZ range of ransboundary species. Showing only species that share > 20 countries](./Figures/Fig3.png)

Due to their large migration patterns and presence in areas beyond national jurisdiction, highly migraroty pelagic fish species are cataloged as straddling and many of them managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species [@CullisSuzuki:2010fi]. We estimate that 81.8% of these species are actually shared within neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas sharing over 40 species with neighboring RFMOs (**Supplemental figure RFMO_MAP**). The effectiveness of RFMOs responsible for managing high migratory stocks has been previously questioned [@CullisSuzuki:2010fi; Lodge:2007bt]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 13 species (40 stocks) suggests that most tuna and billfish stocks are sustainably harvested ($B/B_{MSY} > 1$ and $F/F_{MSY} < 1$) with only three stocks currently overfished [@Pons:2016ep]. Within national waters, we estimated that the catch of transboundary pelagic-oceanic species comes mainly from five Island Nations in the South Pacific from which three (Kiribalti, Solomon and Papua New Guinea) are members of the Pacific Nauru Agreement (Fig 3B). The parties to the Nauru Agreement sustainably manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

@@

<!-- # Materials and Methods -->

<!-- We determine whether or not a species can be considered transboundary. The method is based on a series of criteria to be met and the overlay of species distributions and the spatial boundaries of the world’s EEZs. For the current analysis we work at the species level and adopt the Food and consider transboundary species as those that occur within the Exclusive Economic Zone (EEZ) of neighboring countries [@Munro:2002uf]. All data was scaled to 0.5 degree latitude x 0.5 degree longitude grid cells (*n* = ~180,000 grid cells). -->

<!-- ## Databases on species geographic distributions -->

<!-- To determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs we first extracted all the commercial marine fish or invertebrate species from the *Sea Around Us* database (http://www.seaaroundus.org and determined their current distributions. For this, we used four data sources of species-distributions: (*i*) observational data, (*ii*) an Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (**Table x**). Each source represent a different method of estimating the distribution of a given species and thus, provides a more robust result. Only commercial fished species with data from all four sources were included in the analysis. The final dataset comprised **XXX** species, **XX**% of the reported marine species in the SAU database and **XX**% of the reported taxa of the FAO (**See S1_Data**). -->

<!-- ### *Occurence data* -->

<!-- The occurrence data was collected from five publicly available repositories; Fishbase (http://fishbase.org), the Global Biodiversity Information Facility (GBIF; https://www.gbif.org/), the Ocean Biogeographic Information System (OBIS; https://obis.org/),  the Intergovernmental Oceanographic Comission (IOC; http://ioc-unesco.org), and the International Union for Conservation of Nature (IUCN; https://www.iucn.org/technical-documents/spatial-data) [@Reygondeau:2019uh].   -->

<!-- ### *Distribution models* -->

<!-- In addition to the occurrence data we use two different methods to estimate species distributions, hereafter referred as ENM-Nereus and SDM-SAU. Although they use the same data, the models are structurally different complementing each other and providing robustness to the results. -->

<!-- The ENM-Nereus consists in a multimodel approach based on a Bioclim and a Bososted Regression Tree model [@Thuiller:2009gp], a Maxent model [@Phillips:2006ff], and a Non-Parametric Probabilistic Ecological Niche Model [@Beaugrand:2011fd]. Environmental variables utilized in the models included sea surface temperature, surface pH, surface oxygen concentration, and vertically integrated (0–100 m) net primary production (NPP) [@Asch:2018ca]. The ENM-Nereus employed all of the observational data-sources previously mentioned. Global environmental conditions were obtained from 1970 to 2000 and average from three Earth System Models developed by the Geophysical Fluid Dynamics Laboratory (GFDL- https://www.gfdl.noaa.gov/earth-system-model/), the Institute Pierre Simon Laplace (IPSL- www.icmc.ipsl.fr/), and the Max Planck Institute for Meteorology (MPI- www.mpimet.mpg.de/en/science/models/). See [@Asch:2018ca; @Reygondeau:2019uh] for model details. -->

<!-- The SDM-SAU model follows a five-steps process based on species-specific life history information, rather than environmental variables [@Close:2006ux; @Palomares:wg]. For each commercial fish species, the model first uses the FAO major fishing areas and countries EEZs to determine a broad distribution. It then uses life history information to delimit its range within the FAO fishing area (e.g. thermal preference, depth limits). The range is delimited even further by expert-review polygons and compared with that of AquaMaps [@Kaschner:2016tl], OBIS and GBIF occurrence data. The model then determines a species habitat preference based on the assumptions that the relative abundance of a species is determined by the number of habitats in a grid cell and the distance of the species to each habitat, as well as the importance of the habitat to the species size. Finally, the species equatorial submergence (e.g. the latitudinal region where a species is not seen in between poles) is estimated for each species. See [@Close:2006ux; @Palomares:wg] for model details. -->

<!-- ### *Catch data* -->

<!-- The previous models combine observational data with a series of biotic and abiotic information to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g. ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - i.e., catches  based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in the FAO data available for each country. Catches are also spatially allocated [@Zeller2016]. Thus, we used the Sea Around Us catch reconstruction database from 2005 to 2014 (last ten years of data) as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs. -->

@@

<!-- ## Determine if a species is transboundary -->

<!-- We developed a four-criteria method for determining whether or not a species was transboundary. Only species that meet all criteria were considered as transboundary species and the analysis was done only within the EEZs of coastal states. In some cases, criteria also work as indexes to measure the uncertainty in the analysis as they provided continuous index (e.g. from 0 to 1) rather than a discrete, yes or no, index . All of the analysis was done in the statistical software *R version 3.5.2 (2018-12-20)* with the packages *data.table*[@Packagedatatable:2019uh], *janitor*[@PackagejanitorSim:2018ur], *wesanderson*[@Packagewesanderson:2018vn], *rfishbase*[@PackagerfishbaseR:2019th], *R.matlab*[@PackageRmatlabRe:2018un], *sf*[@PackagesfSimpleF:2018vp], *sp*[@PackagespClasses:2019tn], *tidiverse*[@PackagetidyverseE:2017vq], *tidytext*[@PackagetidytextTe:2019vk], and *zoo*[@PackagezooSInfr:2019tm]. All code is available at https://github.com/jepa/FishForVisa. -->

<!-- ### Criteria 1; Neighboring EEZs -->

<!-- As previously mentioned, we define transboundary species as those marine species that happen within the EEZs of two or more neighboring countries. Hence, the first criteria was that the species would only be transboundary between two neighboring countries, regardless of the species extended distribution. This way, a species such as Atlantic cod (*Gadus morhua*) distributed along the north Atlantic, was only considered transboundary between each of the neighboring nations covering its distribution, rather than the region as a whole. We assume that this would reduce the differences between species and stock as is more likely that neighboring species belong to the same stock, rather than, lets say, Cod in Norway and Canada. However, the analysis here presented was kept at the species level and stocks within countries were not considered. -->

<!-- We define the EEZ boundaries using the SAU shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) that sub-divides EEZs by regions (e.g. Mexico Pacific and Mexico Atlantic) and determine the intersections between polygons using the *sf*-*R* package. When estimating transboundary species, we filter out those shared by the EEZs sub-regions, and when aggregating by country we only counted the species once, if it appeared in more than one sub-region. Species that were present in EEZs that were non-continental territories neighboring other countries were kept (e.g. Argentina and Falkland Islands), but removed when the non-continental territory belonged to the same nation (e.g. Brazil and Fernando de Noronha). -->

<!-- ### Criteria 2; Data agreement -->
